---
title: "Practical1"
---

Setup

```{r}
#| echo: false
library(tibble)
attach(airquality)
airquality
```

Rows with missing values

```{r}
rows <- sort(union(which(is.na(airquality[,1])), which(is.na(airquality[,2]))))
rows <- as_tibble(rows)
rows
```

Summary Statistics

```{r}
#| echo: false

#Tempurature summary statistics 
mean_Temp = mean(Temp)
min_Temp = min(Temp)
max_Temp = max(Temp)
sd_Temp = sd(Temp)

airquality <- na.omit(airquality)

#Ozone Summary statistics
mean_Oz = mean(airquality$Ozone)
min_Oz = min(airquality$Ozone)
max_Oz = max(airquality$Ozone)
sd_Oz = sd(airquality$Ozone)


df <- data.frame(Stats = c("Min", "Mean", "Sd", "Max"), Tempurature = c(min_Temp, mean_Temp, sd_Temp, max_Temp), Ozone = c(min_Oz, mean_Oz, sd_Oz, max_Oz))
df
```

Simple Linear Regression

```{r}
#| echo: false
#| message: false
attach(cars)
```

First principles

```{r}
A <-  matrix(data = cbind(rep(1, 50) ,cars$speed), nrow = 50, ncol = 2)
B <-  cars$dist


#Function that takes response and predictor data frames, estimates the regression parameters under ordinary least squares estiamtion and outputs the results in a list called modmanual
ModelMaker <- function(X,Y){
  
    #Model coefficient estimates
    betas = (solve(t(X)%*%X))%*%(t(X)%*%Y)
    fitted <- X%*%betas #fitted values
    
    #Standard errors
    cov <- solve(t(X)%*%X) #covariance matrix
    SSE <- sum((Y-fitted)**2) #error sum of squares
    n <-  length(Y)
    s <-  sqrt(SSE/(n-2)) #residual standard error
    se1 <- s*sqrt(cov[1]) #standard errors
    se2 <- s*sqrt(cov[4])
    
    #t-values
    t1 <- betas[1]/se1
    t2 <- betas[2]/se2
    
    #p-values
    p1 <- 2*pt(q = t1, df = n-2, lower.tail = TRUE)
    p2 <- 2*pt(q = t2, df = n-2, lower.tail = FALSE)
    
    modmanual <- list(coefficients = round(betas,4), Std.Error = c(round(se1,4),round(se2,4)),     tvalue = c(round(t1,4),round(t2,4)), p = c(round(p1,4),round(p2,4)))

}

ModelMaker(X,Y)

```

The estimates found via first principles are -17.57905 for the intercept coefficient and 3.932409 for the slope coefficient.

Using the lm package, the generated model is:

```{r}
#| echo: false
X2 <- cars$speed
mod <- lm(Y~X2)
summary(mod)
```

We will now test to see if the lm function agrees with our first principal results

```{r}

#Model coefficient differences
round(summary(mod)[["coefficients"]][, "Estimate"],4) == modmanual$coefficients

round(summary(mod)[["coefficients"]][, "Std. Error"],4) == modmanual$Std.Error

round(summary(mod)[["coefficients"]][, "t value"],4) == modmanual$tvalue

round(summary(mod)[["coefficients"]][, "Pr(>|t|)"],4) == modmanual$p


```

Since the differences between the appropriate estimates are negligible, we conclude that thge lm model agrees with the first principal calculations.
